---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
options(digits = 4, width = 120)
```

# queryr: Data validation queries with tidy, stackable output
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

### Installation

Install from GitHub with:

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("epicentre-msf/queryr")
```

### Preliminaries

Load package and examine example dataset `ll`, an epidemiological "linelist".

```{r}
library(queryr)
data(ll)
head(ll)
```

### Example queries

**E.g. 1.** Find observations where `date_exit` is earlier than `date_admit`.

```{r}
query(ll, date_exit < date_admit, cols_base = id:site)
```

Note that by default the columns referenced in the query expression are pivoted
to long-format (`variable1`, `value1`, `variable2`, `value2`, ...), to enable
stacking multiple queries on different variables. The optional `cols_base`
argument can be used to specify additional columns to retain in the output (via
[tidy-selection](https://tidyselect.r-lib.org/reference/select_helpers.html)).

**E.g. 2.** Find any date value in the future using a `.x` selector within the
query expression to represent a set of multiple columns. The columns represented
by `.x` are specified separately via tidy-selection with argument `cols_dotx`.

```{r}
query(ll, .x > Sys.Date(), cols_dotx = starts_with("date"), cols_base = id:site)
```

**E.g. 3.** Find non-valid values of `age_unit`.

```{r}
age_unit_valid <- c("Years", "Months", "Weeks", "Days")
query(ll, !age_unit %in% age_unit_valid, cols_base = id:site)
```

**E.g. 4.** Stack multiple queries on different variables.

Because we'll usually want to retain the same set of 'base columns' across
multiple queries, the argument `cols_base` can be set for an entire session
using option "queryr_cols_base".

```{r}
# set argument `cols_base` for all subsequent queries
options(queryr_cols_base = quote(id:site))

# stack multiple queries
dplyr::bind_rows(
  # ID_01: non-valid patient id
  query(ll, !grepl("^M[0-9]{3}", id), qval = "ID_01"),
  # DATE_01: date in future
  query(ll, .x > Sys.Date(), cols_dotx = starts_with("date"), qval = "DATE_01"),
  # DATE_02: exit before admission
  query(ll, date_exit < date_admit, qval = "DATE_02"),
  # CATEGOR_01: non-valid value of age_unit
  query(ll, !age_unit %in% age_unit_valid, qval = "CATEGOR_01"),
  # LOGIC_01: status 'Confirmed' but lab result not 'Positive'
  query(ll, status == "Confirmed" & !lab_result %in% "Positive", qval = "LOGIC_01")
)
```

