---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
options(digits = 4, width = 120)
```

# queryr: Data validation queries with tidy, stackable output
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R-CMD-check](https://github.com/epicentre-msf/queryr/workflows/R-CMD-check/badge.svg)](https://github.com/epicentre-msf/queryr/actions)
[![Codecov test coverage](https://codecov.io/gh/epicentre-msf/queryr/branch/master/graph/badge.svg)](https://codecov.io/gh/epicentre-msf/queryr?branch=master)

<!-- badges: end -->

### Installation

Install from GitHub with:

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("epicentre-msf/queryr")
```

### Preliminaries

Load package and examine example dataset `ll`, an epidemiological "linelist".

```{r}
library(queryr)
data(ll)
head(ll)
```

### Example queries

**E.g. 1.** Find observations where `date_exit` is earlier than `date_admit`.

```{r}
query(ll, date_exit < date_admit, cols_base = id:site)
```

Note that by default the columns referenced in the query expression are pivoted
to long-format (`variable1`, `value1`, `variable2`, `value2`, ...), to enable
stacking multiple queries on different variables. The optional `cols_base`
argument can be used to specify additional columns to retain in the output (via
[tidy-selection](https://tidyselect.r-lib.org/reference/select_helpers.html)).

**E.g. 2.** Find any date value in the future using a `.x` selector within the
query expression to represent a set of multiple columns. The columns represented
by `.x` are specified separately with argument `cols_dotx`.

```{r}
query(ll, .x > Sys.Date(), cols_dotx = starts_with("date"), cols_base = id:site)
```

**E.g. 3.** Find non-valid values of `lab_result`.

```{r}
lab_result_valid <- c("Positive", "Negative", "Inc.", NA)
query(ll, !lab_result %in% lab_result_valid, cols_base = id:site)
```

**E.g. 4.** Stack multiple queries on different variables.

Because we'll usually want to retain the same set of 'base columns' across
multiple queries, the argument `cols_base` can be set for an entire session
using option "queryr_cols_base".

```{r}
# set argument `cols_base` for all subsequent queries (wrap in quote() if using
# non-standard evaluation rather than a character vector)
options(queryr_cols_base = quote(id:site))

# stack multiple queries
dplyr::bind_rows(
  .id = "query",
  # non-valid patient id
  "IDENT_01" = query(ll, !grepl("^TC\\d{3}", id)),
  # date in future
  "DATES_01" = query(ll, .x > Sys.Date(), cols_dotx = starts_with("date")),
  # exit before admission
  "DATES_02" = query(ll, date_exit < date_admit),
  # non-valid value of age_unit
  "FCTRS_01" = query(ll, !lab_result %in% lab_result_valid),
  # status 'Confirmed' but lab result not 'Positive'
  "LOGIC_01" = query(ll, status == "Confirmed" & !lab_result %in% "Positive")
)
```

### Queries relating to two data frames

For these we'll introduce a second example dataset `sll`, a 'summary' linelist.
Whereas the original linelist `ll` contains all patients presenting at treatment
centres, the summary linelist `sll` contains only patients with status
"Confirmed" or "Probable".

```{r}
data(sll)
head(sll)
```

**E.g. 1.** Find treatment centre IDs that appear in `ll` but not `sll`, and the
corresponding patient status.

```{r}
query2(
  ll,
  sll,
  cols_base1 = c(id, site, status),
  join_type = "anti",
  join_by = c("id" = "tc_id")
)
```

**E.g. 2.** Find confirmed/probable patients in `ll` with different outcomes
listed in `ll` vs `sll`.

```{r}
query2(
  ll,
  sll,
  cond1 = status %in% c("Confirmed", "Probable"),
  cols_base1 = id:site,
  join_type = "inner",
  join_by = c("id" = "tc_id"),
  cond3 = outcome != sll_outcome
)
```
